Article Title

### 2014-01-04 plos.tex
Changes in Introduction, Discussion, Results

We sought the most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix should be most efficiently represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.  Indeed, in cross-validation tests, the covariance matrix estimator with such structure consistently outperformed other regularized estimators. The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.  Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than sample correlations.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neural populations have proven useful to describe their functional connectivity.  Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.  What kinds of correlation effects dominate the population activity in cortical microcircuits? To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.  In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.  Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from the usual correlations. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits and reflect interactions between neurons \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.

@@

\paragraph{Structure  of correlation matrices}
Modern multineuronal recordings allow measuring entire correlation matrices from increasingly large and dense populations of highly interconnected neurons. Multivariate analysis techniques based on large correlation matrices provide information that could not be extracted from an equivalent number of isolated measurements of pairwise correlations.  For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can help isolate various aspects of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.

One reduced representation of the correlation matrix is its low-rank approximation.  Low-rank approximations are particularly suitable for capturing common fluctuations or common external inputs into the circuit and can be extract using principal component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.

Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlation} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}. The partial correlation between a pair of neurons is a measure of their linear association after the effects of the rest of the network have been taken into account \cite{Anderson:2003}. Partial correlations are particularly meaningful when relationships between variables are mostly linear and large fraction of the variables are observed.  In this case, partial correlations approximate conditional dependency.  A lack of conditional dependency between two neurons suggests a lack of direct interaction between them. Networks of partial correlations (sometimes called \emph{association networks}) have been used to infer gene interaction networks \cite{Schafer:2005, Peng:2009} and interactions between brain regions in brain imaging studies \cite{Varoquaux:2012, Ryali:2012}.

\paragraph{Improved estimation performance by regularization}
Imposing some structure on an estimate of the covariance matrix can improve its estimation performance.

@@

In statistics, the technique of deliberately imposing some structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.    To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.  The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.  Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely, without shrinkage. Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).  Yet others can do both target selection and shrinkage toward the target.

Although regularized covariance estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.  
\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972, Friedman:2008}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix (inverse of the covariance matrix) while setting the rest to zero.  Zeros in the off-diagonal elements of the precision matrix indicate zero partial correlations between the corresponding pair of neurons.   Therefore, covariance selection amounts to finding an approximation of the covariance matrix so that a large fraction of neuron pairs are linearly independent of each other when conditioned on the activity of all the other observed cells.

@@

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Conceptual questions and basic approach}
What structures of correlation matrices are most suitable for the description of the multineuronal activity in specific neural circuits and brain states? For example, are correlations in visual cortex under during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?

We pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity. An accurate characterization of the patterns of correlations helps in interpreting them, and to better relate them to the architecture and dynamics of the underlying neuronal networks.

 In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
\item[$C_{\sf sparse}$] -- sparsified precision matrix,  \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
\item[$C_{\sf sparse+latent}$] -- precision matrix composed as the sum of a sparse and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
\end{description}

First, we demonstrated that, in simulations with known correlation structures, regularized estimators with matching structure were generally more efficient than the other estimators.

We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units. Since it led to better estimates, we propose that this structure provides a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.  We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations. In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.

The true correlation matrix is the normalized version of the true covariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates. In the case of noise correlations, $\mu$ depends on the stimulus condition.  Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$,  using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.

@@

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic systems with linear dependencies, these correspond to graphical modes with distinct types of structure  (\figref{02}{row 1}).

@@

For the first regularized estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal contains  variance estimates. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The diagonal target has structure corresponding to the absence of linear associations between the activity of observed neurons (\figref{02}{A}).  If this structure accurately describes recorded data, then strong shrinkage toward $D$ adds little bias while strongly reducing the variability of the estimate. Shrinkage allows for partial commitment to the low-dimensional representation.

The target of the second regularized estimator $C_{\sf factor}$, is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix of rank $d$, and $\Psi$ is the diagonal matrix of independent variances.  The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $L + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between pairs of observed cells (\figref{02}{C}).  This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero.    Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}). The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros off the diagonal. The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.

The fourth estimator, $C_{\sf sparse+latent}$, combines the assumptions of the previous two.  The underlying structure arises from sparse interactions between the observed neurons and several latent units (\figref{02}{D}). The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of rank $d$. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

To evaluate the performance of a covariance matrix estimator $C$, we define a real-valued \emph{loss function} $\loss{C,\Sigma}$ that attains a minimum when $C=\Sigma$.  The loss function quantifies the error of the estimate, \emph{i.e.}~the deviation of the estimate $C$ from truth $\Sigma$.

In this study, we adopted \emph{negative normal log-likelihood loss} defined as
\begin{equation}\label{eq:loss}
    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
\end{equation}
The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$ as
\begin{equation}
    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
\end{equation}

@@

In Figure \ref{fig:02}, row 4 shows the mean excess losses and their standard errors calculated from 30 samples of sizes n=250, 500, 1000, 2000, and 4000 for each estimator and each ground truth. For each estimator, hyperparameters were estimated by nested cross-validation (See Methods for more details.)

As expected, estimators with whose structure matched that of the true model typically outperformed the other estimators. There were two exceptions: First, for small sample sizes, there was insufficient data to reveal the true correlation structure. In this case estimates with simpler targets often outperformed the estimator based on the correct model. Second, the sparse+latent estimator often performed equally well to the sparse estimator even when the data did not have any latent units.  In these cases, it inferred the right number of latent units: zero. In larger models, the falsely identified latent factors are more common.

@@

Then the cross-validation loss is in fact an unbiased estimate of loss.
\begin{equation*}
    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
\end{equation*}

The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses.  With non-additive loss functions, cross-validation loss is a biased estimate of loss, but it is still generally appropriate to be used in lieu of loss.

With empirical data, acquiring a separate testing sample is not practical. Instead, $K$-fold cross-validation is used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).  Then each subset is used as the validation sample with the other $K-1$ serving as a training dataset. The validation losses from each of such \sq{folds} are averaged to produce \emph{cross-validation loss} or \emph{CV-loss} for short.  Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ trained on the remaining $K-1$ subsets. Then cross-validation loss for estimator $C$ is
\begin{equation}\label{eq:cv-loss}
    \ell_C=\frac 1 K \sum\limits_{k=1}^K \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}}
\end{equation}
In the present formulation, $\ell_C$ is known as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier).

Since we only need to compare estimators to each other, we are only interested in \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
\begin{equation}\label{eq:rel-cv-loss}
    \ell_{C,C_{\sf ref}} = \frac 1 K \sum\limits_{k=1}^K \left[
        \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}} -
    \loss{C_{\sf ref}^{\{\setminus k\}},C_{\sf 0}^{\{k\}}} 
\right]
\end{equation}

>>>>

We sought the most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix should be most efficiently represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.  Indeed, in cross-validation tests, the covariance matrix estimator with such structure consistently outperformed other regularized estimators. The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and the difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.  Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than sample correlations.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neural populations have proven useful as descriptions of the functional connectivity in the embedding circuit.  Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.  What kinds of correlation effects dominate the population activity in cortical microcircuits? To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.  In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.  Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from sample correlations. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the functional connectivity in neural circuits \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.

@@

\paragraph{Structure  of correlation matrices}
Modern multineuronal recordings allow measuring entire correlation matrices from increasingly large and dense populations of highly interconnected neurons. Multivariate analysis techniques based on large correlation matrices provide information that could not be extracted from an equivalent number of isolated measurements of pairwise correlations.  For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can isolate an important aspect of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.

One reduced representation of the correlation matrix is its low-rank approximation.  Low-rank approximations are particularly suitable for capturing common fluctuations and common external inputs and can be extracted using principal component analysis, independent component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.

Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlations} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}. The partial correlation between a pair of neurons is a measure of their linear association after the effects of the rest of the network have been taken into account \cite{Anderson:2003}. Partial correlations are particularly meaningful when all effects between the variables are nearly linear and a large fraction of the variables are observed; in this case, partial correlations approximate conditional dependency.  A lack of conditional dependency between two neurons suggests a lack of direct interaction between them. Networks of partial correlations (sometimes called \emph{association networks}) have been used to infer gene interaction networks \cite{Schafer:2005, Peng:2009} and interactions between brain regions in brain imaging studies \cite{Varoquaux:2012, Ryali:2012}.

\paragraph{Improved estimation performance by regularization}
Imposing a structure on an estimate of the covariance matrix can improve its estimation performance.

@@

In statistics, the technique of deliberately imposing a structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.    To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.  The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.  Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely. Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).  Yet others can do both target selection and shrinkage toward the target.

Although regularized covariance matrix estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.  
\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972, Friedman:2008}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix (inverse of the covariance matrix) while setting the rest to zero.  Zeros in the off-diagonal elements of the precision matrix indicate zero partial correlations between the corresponding pair of neurons.   Therefore, covariance selection amounts to finding an approximation of the covariance matrix that maximizes the number of neuron pairs that are linearly independent of each other when conditioned on the activity of all the other observed cells.

@@

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Conceptual questions and basic approach}
What structures of correlation matrices best describe the multineuronal activity in specific circuits and in specific brain states?  More specifically, are correlations in the visual cortex during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?

To address these questions, we evaluated five covariance matrix estimators based on different assumptions about the underlying correlation structure:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix. 
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis.
\item[$C_{\sf sparse}$] --  sparse partial correlations, \emph{i.e.}\;covariance selection.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations with several inferred latent units.
\end{description}

First, we demonstrated that, in simulations with known true correlation structures, estimators that imposed a matching structure were generally more efficient than the other estimators.  The estimators’ relative performance was evaluated by cross-validation.   The logic of this study relies on the converse of this relationship: the covariance matrix estimator that is found to be most efficient on empirical data is proposed to provide a better picture of the correlation structure than other estimators.  Caveats for this kind of inference are reviewed in the Discussion.

Therefore, we performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with fast 3D random-access two-photon imaging of calcium signals. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units.  We examined the structure of the sparse partial correlations inferred by the estimate with respect to the neurons’ geometrical positions and preferred orientations and found that stronger dependencies than those of sample correlations. In Discussion, we outline possible approaches to further corroborate the relationship between the structure of the functional connectivity revealed by such methods to the anatomical organization of the circuit.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The icovariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates. In the case of noise correlations, $\mu$ depends on the stimulus condition.  Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$,  using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.

@@

We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic systems with linear dependencies, these correspond to graphical modes with distinct types of structure  (\figref{02}{row 1}).

@@

For estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal contains  variance estimates. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure of target $D$ favors the absence of linear associations between the activity of observed neurons (\figref{02}{A}).  If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ should outperform the other estimators.

The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix of rank $d$, and $\Psi$ is the diagonal matrix of independent variances.  The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $L + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

Estimator $C_{\sf sparse}$ is based on the assumption that correlations result from direct linear associations between pairs of observed neurons and that only a fraction of neuron pairs interact (\figref{02}{C}).  This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero. The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.  The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.

Estimator $C_{\sf sparse+latent}$ allow the assumptions of both $C_{\sf factor}$ and $C_{\sf sparse}$.  Its correlation structure can be explained by linear effects exerted between some pairs of observed neurons and several latent units (\figref{02}{D}). The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of rank $d$. The estimator has two hyperparameters: one to regulate the number of latent units, the other to regulate the sparsity of matrix $S$.

@@

An evaluation of the performance of a covariance matrix estimator $C$ requires a real-valued \emph{loss function} $\loss{C,\Sigma}$. The loss loss function quantifies the deviation of the estimate $C$ from truth $\Sigma$ and attains its minimum when $C=\Sigma$.

In this study, we adopted the \emph{normal loss} defined as
\begin{equation}\label{eq:loss}
    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
\end{equation}

The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelih\emph{normal log-likelihood loss} defined as ood function $L(\Sigma|C_{\sf 0})$ as
\begin{equation}
    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
\end{equation}
Although the normal log-likelihood loss and the normal log-likelihood appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from the estimate $C$ to the truth $\Sigma$.

@@

We computed the excess losses for 30 samples for each sample size $n=250$, 500, 10000, 2000, and 4000 for each of the five estimators and for each ground truth (\figref{02}{\,row 5}). For each estimator, hyperparameters were estimated by nested cross-validation (See Methods.).  As expected, estimators whose structure matched that of the true model typically outperformed the other estimators. There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with the simpler structure, $C_{\sf diag}$ and $C_{\sf factor}$ outperformed the estimators with matching structure (\figref{02}{\,columns C and D, rows 5 and 6}).  Second, $C_{\sf sparse+latent}$ often performed equally well to the sparse estimator even when the ground truth did not have any latent units.  In these cases, it inferred the right number of latent units, zero, and estimated optimally.

@@

Then the cross-validation loss is in fact an unbiased estimate of loss.
\begin{equation*}
    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
\end{equation*}

The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses.  With non-additive loss functions, validation loss is a biased estimate of loss, but it is still generally appropriate to be used in lieu of loss.

With empirical data, acquiring a separate testing sample is not practical. Instead, $K$-fold cross-validation is used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).  Then each subset is used as the validation sample with the other $K-1$ serving as the training sample. The validation losses from each of such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.  Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ trained on the remaining $K-1$ subsets. Then cross-validation loss for estimator $C$ is
\begin{equation}\label{eq:cv-loss}
    \ell_C=\frac 1 K \sum\limits_{k=1}^K \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}}
\end{equation}
In this formulation, $\ell_C$ can be recognized as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier). However, the present framework is more general, allowing for other loss functions to be used.

Since we only need to compare estimators to each other, we are only interested in the \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
\begin{equation}\label{eq:rel-cv-loss}
    \Delta\ell_{C,C_{\sf ref}} = \ell_{C}-\ell_{C_{\sf ref}}
\end{equation}

In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (\figref{02}{Row 6}). For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.

These simulations demonstrated that, with sufficiently large sample sizes, the most efficient among several regularized estimators could be used to infer the most likely type of correlation structure present in the data.
